\textbf{Real-time guitar transcription using\\Fourier transform based methods}\\A pitch estimation framework and overtone sieve algorithm

### 2021-11-30 paper.tex
small fixes

Most modern instruments use the twelve-tone equal temperament tuning system (12-TET). This tuning system divides an octave, which the distance between a frequency and its double, into twelve equal parts on the logarithmic scale. In other words, the ratio between two contiguous notes is $\sqrt[12]{2}$. The distance between two notes can further be divided on a logarithmic scale into 100 cents. The number of cents between two frequencies can be calculated as $1200 * \log_2(\frac{f_1}{f_2})$. 12-TET is tuned relative to a standard pitch. In modern music, we define the $\text{A}_4$ to be 440 Hz.

When we refer to real-time in this paper, we use the definition where a system has finish computations within and produce an answer within a real-time constraint. In this paper, we choose a constraint of 6 milliseconds. We elaborate on that decision in Section~\ref{sec:lat}.

@@

\subsection{Harmonic overtone sieve}  \label{sub:impnoteset}
An overview of the monophonic is given in Algorithm~\ref{alg:mono}. Here, we iterate over all found peaks and check if every higher peak can be an overtone of the considered peak. After counting the number of overtones for every peak, we can pick the peak with the most overtones.
\input{mono.pc}
\input{poly.pc}

@@

\section{Conclusions}
% The used audio equipment has a large effect on the results and in turn effects what parameters work best for the system.
In this paper, we presented and implemented a Fourier transform based AMT system and tested if these methods are feasible for real-time transcription. We divided the task in three components which can all be separately interchanged to facility specialized research in one of the components. The components are: obtaining the frequency domain representation, spectral peak picking and note selection. As shown in Section~\ref{sub:algspeed}, these tasks can be performed well within our real-time constraint of 6 ms. However, when using Fourier transform based methods, we need large frame sizes to get the required accuracy to discern lower pitched notes. Given the maximum sampling rate of widespread audio interfaces, it takes too long to fill a frame. The current implementation requires frame sizes of 16384 samples for the necessary resolution, which equates to frame lengths of 85 ms. As described in Section~\ref{sec:future}, we might still be able to use Fourier transform based methods for real-time transcription. However, other transforms such as CQT or wavelet transform might be better suited for real-time transcription, as we can still spare much computation time to lower the frame-wait time.

@@

The Fourier transform requires too many samples in a frame to remain within the real-time constraint while still providing the necessary frequency resolution in the fundamental frequencies. It might be possible to alleviate this problem with other transforms, such as the constant-Q transform or the wavelet transform, but those transforms may have to solve other problems. As mentioned before, the low resolution does not have to be a problem, as less resolution is needed when discerning higher pitched peaks due to the exponential nature of octaves. Currently, we do not detect enough overtones, but some testing showed us we can detect many overtones if we amplify the signal. This has to be done carefully, as artifacts may be introduced in the signal. Optimally amplifying of a recording is trivial, however, it is trickier in real-time. Using amplified signals would require recalibration of the peak picking parameters or even require a new algorithm in general.

>>>>

Most modern instruments use the twelve-tone equal temperament tuning system (12-TET). This tuning system divides an octave, which is the distance between a frequency and its double, into twelve equal parts on the logarithmic scale. In other words, the ratio between two contiguous notes is $\sqrt[12]{2}$. The distance between two notes can further be divided on a logarithmic scale into 100 cents. The number of cents between two frequencies can be calculated as $1200 * \log_2(\frac{f_1}{f_2})$. 12-TET is tuned relative to a standard pitch. In modern music, we define the $\text{A}_4$ to be 440 Hz.

When we refer to real-time in this paper, we use the definition where a system has finish computations and produce an answer within a real-time constraint. In this paper, we choose a constraint of 6 milliseconds. We elaborate on that decision in Section~\ref{sec:lat}.

@@

\subsection{Harmonic overtone sieve}  \label{sub:impnoteset}
An overview of the monophonic algorithm is given in Algorithm~\ref{alg:mono}. Here, we iterate over all found peaks and check if any higher peak can be an overtone of the considered peak. After counting the number of overtones for every peak, we can pick the peak with the most overtones.
\input{mono.pc}
\input{poly.pc}

@@

\section{Conclusions}
% The used audio equipment has a large effect on the results and in turn effects what parameters work best for the system.
In this paper, we presented and implemented a Fourier transform based AMT system and tested if these methods are feasible for real-time transcription. We divided the task in three components which can all be separately interchanged to facilitate specialized research in one of the components. The components are: obtaining the frequency domain representation, spectral peak picking and note selection. As shown in Section~\ref{sub:algspeed}, these tasks can be performed well within our real-time constraint of 6 ms. However, when using Fourier transform based methods, we need large frame sizes to get the required accuracy to discern lower pitched notes. Given the maximum sampling rate of widespread audio interfaces, it takes too long to fill a frame. The current implementation requires frame sizes of 16384 samples for the necessary resolution, which equates to frame lengths of 85 ms. As described in Section~\ref{sec:future}, we might still be able to use Fourier transform based methods for real-time transcription. However, other transforms such as CQT or wavelet transform might be better suited for real-time transcription, as we can still spare much computation time to lower the frame-wait time.

@@

The Fourier transform requires too many samples in a frame to remain within the real-time constraint while still providing the necessary frequency resolution in the fundamental frequencies. It might be possible to alleviate this problem with other transforms, such as the constant-Q transform or the wavelet transform, but those transforms might introduce other problems as they are less researched. As mentioned before, the low resolution does not have to be a problem, as less resolution is needed when discerning higher pitched peaks due to the exponential nature of octaves. Currently, we do not detect enough overtones, but some testing showed us we can detect many overtones if we amplify the signal. This has to be done carefully, as artifacts may be introduced in the signal. Optimally amplifying of a recording is trivial, however, it is trickier in real-time. Using amplified signals would require recalibration of the peak picking parameters or even require a new algorithm in general.
